Social Conditions of Fear

### 2023-11-29 scripts/manuscript.qmd
github merge test
test 2
tables and figures progress from Nov 28 evening

Race/ ethnicity is an important control for our models. First, we sought to create a combined race variable from the variables "race" and "hispanic." After looking at the data, we saw that the options for the race variable were "Black, White, Other." This does not provide the control we need. So, we opted to use the "racerank1" variable, because participants were able to select from 16 different options: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic. We factored and recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other. We understand that these categories are problematic and contentious, however they are the standard with quantitative data.

@@

>>>>

Informed by theory, race/ ethnicity is an important covariate for our models. First, we sought to create a combined race variable from the variables "race" and "hispanic." After looking at the data, we saw that the options for the race variable were "Black, White, Other." This does not provide the control we need. So, we opted to use the "racerank1" variable, because participants were able to select from 16 different options: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic. We factored and recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other. We understand that these categories are problematic and contentious, however they are the standard with quantitative data.

Race/ ethnicity is an important control for our models. First, we sought to create a combined race variable from the variables “race” and “hispanic.” After looking at the data, we saw that the options for the race variable were “Black, White, Other.” This does not provide the control we need. So, we opted to use the “racerank1” variable, because participants were able to select from 16 different options: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic. We factored and recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other. We understand that these categories are problematic and contentious, however they are the standard with quantitative data.

@@

# Tables and Figures

```{r}
view(gss_subset)

skimr::skim(gss_subset)
```

```{r}
gss_subset %>% 
  select(fear, sexnow1, racecombo1, degree) %>% 
  gtsummary::tbl_summary()
```

```{r}
gss_subset %>% 
  tabyl(fear, sexnow1) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns(position = "front") %>% 
  adorn_title(
    row_name = "Fear at Night",
    col_name = "Gender",
    placement = "combined"
  ) %>% 
  flextable::flextable() %>% 
  flextable::autofit()
```

```{r}
gss_subset %>% 
  tabyl(fear, racecombo) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns(position = "front") %>% 
  adorn_title(
    row_name = "Fear at Night",
    col_name = "Race",
    placement = "combined"
  ) %>% 
  flextable::flextable() %>% 
  flextable::autofit()
```

```{r}
gss_subset %>% 
  ggplot(aes(x = sexnow1, y = after_stat(prop), group = fear, 
             fill = fear)) +
  geom_bar(position = "dodge") + 
  scale_y_continuous(labels = scales::percent) + 
  labs(title = "Social Fear by Gender",
       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
       x = "Gender",
       y = NULL,
       fill = "Fear"
       ) +
  theme_bw()
```

```{r}
gss_subset %>% 
  count(sexnow1, fear) %>% 
  ggplot() +
  geom_col(
    mapping = aes(x = fear,
                  fill = sexnow1,
                  y = n)
  ) +
  coord_flip() +
  labs(title = "Social Fear by Gender",
       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
       x = "Count",
       y = "Fear",
       )+
  theme_bw()
```

```{r}
gss_subset %>% 
  filter(!is.na(racecombo)) %>% 
  filter(racecombo != "Other") %>% 
  ggplot(aes(x = racecombo, y = after_stat(prop), group = fear, 
             fill = fear)) +
  geom_bar(position = "dodge") + 
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Social Fear by Race",
       caption = "Question: Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?",
       x = "Race",
       y = NULL,
       fill = "Fear"
       ) +
  theme_bw()
```
